---
title: "Week_5_lab"
author: "CHAKAR"
date: "2026-07-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Task: Filter on Christchurch data only. Use only this filtered dataset from now on. Using 'june_2026_listings.csv' first

```{r june_2026_listings}

library(tidyverse)
jun_26 <- read_csv('week_5_datasets/june_2026_listings.csv')

print("The first 10 rows are:")
print(head(jun_26, 10))

print("Summary of the dataset is:")
print(summary(jun_26))
```

```{r}
# Check neighbourhood_group column has Christchurch city - list all unique cities in that column
unique(jun_26$neighbourhood_group)
```

```{r}
# Filter out so data only has Christchurch city rows and adding the month+year column
chch_listings <- jun_26 %>%
  filter(neighbourhood_group == 'Christchurch City') %>%
  mutate(month_year = '2026-06')
head(chch_listings)
```

The filtered on Christchurch city data is named: chch_listings

Filtering all the data sets to Christchurch city and adding month + year column using a pipeline

```{r}
load_chch_listings <- function(filename, month_year) {
  read_csv(paste0('week_5_datasets/', filename)) %>%
    filter(neighbourhood_group == 'Christchurch City') %>%
    mutate(month_year = month_year)
}
```

Running the pipeline for the data

```{r}
chch_oct25 <- load_chch_listings('october_2025_listings.csv', '2025-10')
chch_nov25 <- load_chch_listings('november_2025_listings.csv', '2025-11')
chch_dec25 <- load_chch_listings('december_2025_listings.csv', '2025-12')
chch_jan26 <- load_chch_listings('january_2026_listings.csv', '2026-01')
chch_feb26 <- load_chch_listings('february_2026_listings.csv', '2026-02')
chch_mar26 <- load_chch_listings('march_2026_listings.csv', '2026-03')
chch_apr26 <- load_chch_listings('april_2026_listings.csv', '2026-04')
chch_may26 <- load_chch_listings('may_2026_listings.csv', '2026-05')
chch_jun26 <- load_chch_listings('june_2026_listings.csv', '2026-06')
```

combining datasets into full data set for Oct 2025 to June 2026 for Christchurch

```{r}
all_chch <- bind_rows(chch_oct25, chch_nov25, chch_dec25, chch_jan26, chch_feb26, chch_mar26, chch_apr26, chch_may26, chch_jun26)

```

Summary stats

```{r}
all_chch %>%
  select(-id, -host_id, -latitude, -longitude) %>%
  summarise(across(where(is.numeric),
                    list(min = ~min(.x, na.rm = TRUE),
                         max = ~max(.x, na.rm = TRUE),
                         mean = ~mean(.x, na.rm = TRUE),
                         sd = ~sd(.x, na.rm = TRUE))))
```

```{r}
table(all_chch$room_type)
table(all_chch$neighbourhood)
```

```{r}
colSums(is.na(all_chch))
```

Storing the combined Christchurch dataset

```{r}
write_csv(all_chch, 'week_5_datasets/all_chch_listings.csv')
```
