---
title: "week_8_lab_airbnb"
author: "CHAKAR"
date: "2026-09-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Week 8 instruction for Airbnb file:
All_chch listings (separate datasets from oct 2025 to jun 2026 combined, but only for chch city).
clean the dataset -> ultimate aim is going to be comparing rental prices and number of available properties with tenancy dataset

```{r}
# load data as "data"
library(tidyverse)

data <- read_csv("week_8_datasets/all_chch_listings.csv")

# summary statistics for all columns
print(summary(data))
```
data has: Rows: 28795 Columns: 19

```{r}
# datatype
print(glimpse(data))
```

```{r}
# number of missing NA values
print(colSums(is.na(data)))
```

```{r}
# plot price distribution
ggplot(data, aes(x = price)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white") +
  theme_minimal() +
  labs(title = "Distribution of Rental Prices", x = "Price", y = "Count")
```

Instruction from week 5 lab - so probably need days_since_review column - added that column in
```{r}
# days since last reivew (calculate the time difference between scrape/publish date) then plot it
data <- data %>%
  mutate(
    scrape_date = ymd(paste0(month_year, "-01")), # Converts "YYYY-MM" to date # scrap date set to 1st of each month
    last_review_date = ymd(last_review),
    days_since_review = as.numeric(scrape_date - last_review_date)
  )
head(data)

ggplot(data, aes(x = days_since_review)) +
  geom_histogram(bins = 50, fill = "forestgreen", color = "white") +
  theme_minimal() +
  labs(title = "Days Since Last Review", x = "Days", y = "Count")

# find number of negative days
negative_rows <- data %>% 
  filter(days_since_review < 0)
negative_rows
```
Summary statistics: 
price: median price is $205, but the max is $43,654 (a massive extreme outlier? error?).
Price histogram: right skewed, due to extreme price outliers

license: entirely NA (28,795 missing values) -> consider dropping if doesn't need this column

Missing values: 
price has 10,667 missing values: how to handle? or filter them out?

last_review and reviews_per_month both have 2,627 missing values

days_since_review has 9418 negative values. Days since last reivew histogram: Most reviews happened very recently (clumped near 0 days), which makes sense for active listings, but there are long-tail stale listings stretching out past 1,000+ days.
For month_year column, change values to include the day the scrape date was done (from the website).
```{r}
data_1 <- data %>%
  mutate( # create/modify existing columns
    scrape_date = case_when(
      month_year == "2025-10" ~ ymd("2025-10-05"), #f month_year is "2025-10", the code assigns the scrape date October 5, 2025, ymd() converts the text "2025-10-05" into an R date
      month_year == "2025-11" ~ ymd("2025-11-07"),
      month_year == "2025-12" ~ ymd("2025-12-11"),
      month_year == "2026-01" ~ ymd("2026-01-16"),
      month_year == "2026-02" ~ ymd("2026-02-13"),
      month_year == "2026-03" ~ ymd("2026-03-17"),
      month_year == "2026-04" ~ ymd("2026-04-16"),
      month_year == "2026-05" ~ ymd("2026-05-23"),
      month_year == "2026-06" ~ ymd("2026-06-19"),
      TRUE ~ NA_Date_ # if months isn't one of the months listed, leave the scrape date blank
    ),
    last_review_date = ymd(last_review),
    days_since_review = as.numeric(scrape_date - last_review_date)
  )

# Check how many rows in days_since_review still has negative values
negative_rows <- data_1 %>% 
  filter(days_since_review < 0)
print(negative_rows)
```

Even after matching the scrape dates with the exact publication times, 1745 negative values remain in days_since_review column.
(So these reviews were posted in the days right after each monthly data collection, counted as active listing. So leave these negative days for now)


DEAL with price column first -> extreme outliers + 10667 missing values
Check Price inconsistencies
```{r}
# Check the data type of the price column
class(data_1$price)

# 3. Check how many missing values (NAs) exist in the price column
sum(is.na(data_1$price))
```
See if 10667 missing prices are random or linked to specific months
```{r}
# Check missing prices grouped by each month scrape
missing_by_month <- data_1 %>%
  group_by(month_year) %>%
  summarise(
    total_listings = n(),
    missing_prices = sum(is.na(price)),
    pct_missing = (missing_prices / total_listings) * 100
  )

print(missing_by_month)
```

100% price missing values in 2025-12, 2026-01, 2026-02 months

checking whether a listing has only one unique price across all its available months. If it does, it's stable, can use that price to fill in December, January, and February.
```{r}
# Check if listing names change across months for the same ID
name_check <- data_1 %>%
  filter(!is.na(name)) %>%
  group_by(id) %>%
  summarise(unique_names = n_distinct(name))

# See how many listings ever changed their title
table(name_check$unique_names > 1)
```
FALSE (3,535 listings): For the vast majority of properties, the host never changed the listing title across the entire time period you observed. Their title stayed 100% identical.

TRUE (582 listings): For 582 properties, the host did change their listing title at least once during your dataset
```{r}
# 1. Check price consistency for each listing across available months
price_consistency <- data_1 %>%
  filter(!is.na(price)) %>% # throw away rows where price is missing
  group_by(id) %>% # group all rows of individual listings based on id
  summarise( # each listing id becomes a single row
    unique_prices = n_distinct(price), # Counts how many different prices a listing had
    consistent_price = first(price),   # Grabs the first price it sees for that listing
    months_observed = n() # count how many total valid monthly data points exist for that specific listing in the dataset
  )
  
# Impute prices ONLY for listings that have a proven stable price history
# See how many listings have stable (unchanging) prices vs changing prices
table(price_consistency$unique_prices)

# Create a lookup table of ONLY the strictly stable listings
stable_price_lookup <- price_consistency %>%
  filter(unique_prices == 1) %>%
  select(id, consistent_price) 

# Impute prices safely using ifelse() to prevent type mismatch crashes
data_2 <- data_1 %>%
  left_join(stable_price_lookup, by = "id") %>%
  mutate(
    # If price is missing AND it matches the target months, pull from consistent_price
    price = ifelse(
      is.na(price) & month_year %in% c("2025-12", "2026-01", "2026-02"), 
      consistent_price, 
      price
    )
  ) %>%
  select(-consistent_price) # clean up temporary lookup column

# Check how many missing prices remain
print(sum(is.na(data_2$price)))
```
So 509 prices were filled as they were the same across every single other months they appeared in (id)

Still has 10158 missing prices


Manage price outliers/extreme values
Inspect the extreme rows
```{r}
# Check the top 20 most expensive listings using summary-file columns
top_expensive <- data_2 %>%
  filter(!is.na(price)) %>%
  arrange(desc(price)) %>%
  select(id, name, price, room_type, minimum_nights) %>%
  head(20)

print(top_expensive)
```
Rows 1-8 -> likely error (checking latitude/altitude), doesn't make sense to have that much a night
Rows 10-15 -> also likely error. priced >5000 per night in a private room doesn't make sense
Rows 16-20.. maybe? entire home

Keep rows 1-15 will cause statistical bias into analysis. It distorts the mean and standard deviation
DECISION: Filter these 15 rows 
(note: filter() keeps every row where the condition evaluates to TRUE, and throws away anything that evaluates to FALSE)
```{r}
# Clean the dataset by removing anything $5,000 or higher (while keeping NAs. Otherwise R will remove NAs as well)
data_3 <- data_2 %>%
  filter(price < 5000 | is.na(price))

# Check how many rows were dropped
nrow(data_2) - nrow(data_3)
```
```{r}
# 1. Check the new maximum price to ensure the errors are truly gone
max(data_3$price, na.rm = TRUE)

# 2. Calculate clean summary statistics for prices
price_summary <- data_3 %>%
  filter(!is.na(price)) %>%
  summarise(
    mean_price = mean(price),
    median_price = median(price),
    min_price = min(price),
    max_price = max(price)
  )

print(price_summary)
```
What to do with remaining PRICE NAs? - leave them as NAs for now as not sure what is happening downward

License 100% NAs - no universal registration rules so hosts didn't have to register for official registration code to list their property
-> DECISION DROP icense column
```{r}
# Drop the useless license column
df_cleaned <- data_3 %>%
  select(-any_of("license"))

# check columns
colnames(df_cleaned)
```
check if reviews_per_month and last_review are related
```{r}
# Check if the exact same rows are missing values for both columns
identical(is.na(df_cleaned$reviews_per_month), is.na(df_cleaned$last_review))
```
if property has 0 reviews, both reviews_per_month, and last_review are missing. 
```{r}
# 1. Check if there are any literal 0 values already in reviews_per_month
sum(df_cleaned$reviews_per_month == 0, na.rm = TRUE)

# 2. Check if there are any non-NA values in last_review when reviews_per_month is NA (or vice versa)
df_cleaned %>%
  filter(is.na(reviews_per_month) != is.na(last_review)) %>%
  select(id, name, reviews_per_month, last_review)
```
There are zero listings.. every single listing either has a positive review or is missing (NA). So reviews_per_month and last_review are 100% structurally locked

DECISION: takes every missing value (NA) in the reviews_per_month column and turns it into a literal 0
```{r}
# Finalize the review columns treatment based on structural verification
df_cleaned <- df_cleaned %>%
  mutate(
    # Recode NA to 0 for numerical analysis (since we proved 0 doesn't otherwise exist)
    reviews_per_month = replace_na(reviews_per_month, 0)
    # last_review remains NA, perfectly justified by the zero-review status
  )

# 1. Check how many NA values remain in reviews_per_month (should be 0)
sum(is.na(df_cleaned$reviews_per_month))

# 2. Check how many 0s are now in reviews_per_month (should match your previous missing count, 2627)
sum(df_cleaned$reviews_per_month == 0)
```
```{r}
# check columns
colSums(is.na(df_cleaned))
```
The Review-Related Columns (2,619 NAs)
last_review: 2,619
last_review_date: 2,619
days_since_review: 2,619
These three columns are locked together
Decision to leave them as NA since can't calculate "days since a review" if there has never been a review.

Price column: 10158 NAs

minimum_nights: 37 NAs

TO DECIDE: MINIMUM_NIGHTS NAs ?impute with median : min-1, max-365, median-1, mean - 1.788

COLUMNS to drop:
neighbourhood group only has 1 uniqe value so can't help compare anything
```{r}
# Check is the right column
head(df_cleaned$neighbourhood_group)
```
DECISION: DROP neighbourhood_group column
```{r}
# Drop the zero-variance column (replace with the exact column name, e.g., neighbourhood_group)
df_cleaned <- df_cleaned %>%
  select(-neighbourhood_group)
colnames(df_cleaned)
```
```{r}
#check to see if last_review and last_review_date contain the exact same data
# Check if both columns are identical element-by-element
identical(df_cleaned$last_review, df_cleaned$last_review_date)
```
DECISIONS:
Drop: 
host_name : irrelevant to analysis, and is personal identifiable information
Drop last_review_date: Both last_review and last_Review_Date are the same

```{r}
df_cleaned <- df_cleaned %>%
  select(-c(host_name, last_review_date))
colnames(df_cleaned)
```
```{r}
#Check if any listing has more LTM reviews than total reviews (should output 0 rows)
df_cleaned %>%
  filter(number_of_reviews_ltm > number_of_reviews)
```
Check duplicates
```{r}
# Count how many fully duplicated rows exist
sum(duplicated(df_cleaned))

# Count how many duplicate IDs exist
sum(duplicated(df_cleaned$id))

#See which specific IDs are duplicated and how many times
df_cleaned %>%
  count(id) %>%
  filter(n > 1)

#check if these duplicate IDs are just the same property showing up in different months
# Pick one of the duplicated IDs from your output and see its monthly breakdown
df_cleaned %>% 
  filter(id == 204050) %>% # replace 204050 with one of your actual duplicated IDs
  select(id, month_year, price, availability_365)
```
The same listing IDs appear multiple times because Inside Airbnb scraped them across different months, capturing how their prices and availability dynamically changed over time.

This means they are not accidental duplicate errors. They are time-series records.

```{r}
# 1. Quick check of column types (ensure prices/lat/long are numeric, dates are proper, etc.)
glimpse(df_cleaned)

# 2. Confirm total rows remaining
nrow(df_cleaned)
```
Summary:
1.price column
- Initially 10667 missing values, with 100% missing in 2025-12, 2026-01, 2026-02 months.
- Imputed price for listings that had the same price history across all other months (509 prices were filled).
- (Outlier Removal) Removed price column top 15 rows, priced >5000 as they are very likely errors. Total number of rows in price column before cleaning: 28795, after cleaning: 28780.
- Question: what to do with the remaining 10158 missing values in price column? The column is still right_skewed.

2. week 5 instructed a days_since_review column. So that column added in. Initially scrape_date column used YYYY-MM-01 but had 9418 Negative values in the days_since_review column. So scrape_date column added the scrap "day" got from the website, and recalculated days_since_review column.
Question: still has 1745 negative days (reviews likely posted in the days a). Do we need this column and what to do with this negative days?

3. Removed license column (100% NAs)
4. Removed neighbourhood_group because it only had "Christchurch_city" value so can't used to compare with anything.
5. Removed host_name: has personal identifiable information, and appeared irrelevant to analysis.
6. Removed last_review_date, because both last_review and last_review_date have exactly same values.

7. turned 2619 missing values (NA) in the reviews_per_month column and turned it into numerical 0. Because if property has 0 reviews, both reviews_per_month and last_review columns have missing values (NA). (Every listing either has a positive review >0 count, or is missing (NA)). Question - But left last_review NAs as NAs for now since 0 can mean something in that column.

8.minimum_nights column has 37 NAs. 
Question - impute it with median (min-1, max-365, median-1, mean - 1.788)? or leave it as NAs


