---
title: "Week_5_lab"
author: "CHAKAR"
date: "2026-07-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Task: Filter on Christchurch data only. Use only this filtered dataset from now on. Using 'june_2026_listings.csv' first

```{r june_2026_listings}

library(tidyverse)
jun_26 <- read_csv('week_5_datasets/june_2026_listings.csv')

print("The first 10 rows are:")
print(head(jun_26, 10))

print("Summary of the dataset is:")
print(summary(jun_26))
```

```{r}
# Check neighbourhood_group column has Christchurch city - list all unique cities in that column
unique(jun_26$neighbourhood_group)
```

```{r}
# Filter out so data only has Christchurch city rows and adding the month+year column
chch_listings <- jun_26 %>%
  filter(neighbourhood_group == 'Christchurch City') %>%
  mutate(month_year = '2026-06')
head(chch_listings)
```

The filtered on Christchurch city data is named: chch_listings

Filtering all the data sets to Christchurch city and adding month + year column using a pipeline
```{r}
load_chch_listings <- function(filename, month_year) {
  read_csv(paste0('week_5_datasets/', filename)) %>%
    filter(neighbourhood_group == 'Christchurch City') %>%
    mutate(month_year = month_year)
}
```

Running the pipeline for the data
```{r}
chch_oct25 <- load_chch_listings('october_2025_listings.csv', '2025-10')
chch_nov25 <- load_chch_listings('november_2025_listings.csv', '2025-11')
chch_dec25 <- load_chch_listings('december_2025_listings.csv', '2025-12')
chch_jan26 <- load_chch_listings('january_2026_listings.csv', '2026-01')
chch_feb26 <- load_chch_listings('february_2026_listings.csv', '2026-02')
chch_mar26 <- load_chch_listings('march_2026_listings.csv', '2026-03')
chch_apr26 <- load_chch_listings('april_2026_listings.csv', '2026-04')
chch_may26 <- load_chch_listings('may_2026_listings.csv', '2026-05')
chch_jun26 <- load_chch_listings('june_2026_listings.csv', '2026-06')
```

combining datasets into full data set for Oct 2025 to June 2026 for Christchurch
```{r}
all_chch <- bind_rows(chch_oct25, chch_nov25, chch_dec25, chch_jan26, chch_feb26, chch_mar26, chch_apr26, chch_may26, chch_jun26)

```

Combined dataset called: all_chch

Summary stats
```{r}
# min + max + mean + std of columns
all_chch %>%
  select(-id, -host_id, -latitude, -longitude) %>%
  summarise(across(where(is.numeric),
                    list(min = ~min(.x, na.rm = TRUE),
                         max = ~max(.x, na.rm = TRUE),
                         mean = ~mean(.x, na.rm = TRUE),
                         sd = ~sd(.x, na.rm = TRUE))))
```

```{r}
# frequency table counting how many times unique category appears in room_type coumn and neighbourhood column

table(all_chch$room_type)
table(all_chch$neighbourhood)
```

```{r}
# Calculate number of missing values per column
colSums(is.na(all_chch))
```

Storing the combined Christchurch dataset named: all_chch_listings.csv
```{r}
write_csv(all_chch, 'week_5_datasets/all_chch_listings.csv')
```


``` {r all_chch_listings.csv tasks}
chch_file <- read_csv('week_5_datasets/all_chch_listings.csv')
head(chch_file, 10) # check and read first 10 rows of the chch_file
```
Task Reproduce the workflow and the plots in R 
Plot the distribution of price as a histogram for christchurch city

(also said to plot for all NZ but data has gotten rid of NZ?)
(also said special task to plot both in the same plot but..?)

```{r plot distribution of price as a histogram for Christchurch city}
hist(chch_file$price,
     main = "Distribution of Price for Christchurch City", 
     xlab = "Price",
     ylab = "Number of listings",
     col = "blue",
     border = "white")
```
```{r check price vs frequency}

# Check the price vs frequency in each bin in a table form
# Create 5000 block/bins from 0 to 45000
price_bins <- cut(chch_file$price, breaks =seq(0, 45000, by = 5000))
table(price_bins)
```

1 listing priced between 20000-25000, and 8 priced at 40000-45000

```{r check outliers}

outliers <- chch_file %>%
  filter(price >20000)
view(outliers)
```

To check the accommodation priced at $22971/night. Checking its latitude (-43.52059) and longitude(172.6496) showed on google map to be at address 333 Bealey Avenue, chch central. 
Checked accommodation on google map priced at $43188/night and $43178/night, also appeared unlikely to worth that much.
These 9 values likely error, to drop these 9 extreme outlier values

```{r dropping 9 price outliers}

chch_clean <- chch_file %>%
  filter(price <=1500) #filter to only include accommodation with price <=$1500

hist(chch_clean$price,
     main="Distribution of Price for Christchurch city",
     xlab = "Price",
     ylab = "Number of listings",
     col="blue",
     border="white")
```


Task- Calculate how long ago the last review was in days (difference between scrape/publish date and last review date). Plot the distribution of this new feature.

(go back to use the full chch dataset variable name: chch_file, as reviews may still be correct)

```{r review days}
# check last_review data type
class(chch_file$last_review)

# visualise what the dates look like on last_review - first 5 rows
head(chch_file$last_review)
```
```{r review days}

# Using last publish date on airbnb listings website, which is 19 Jun 2026

publish_date <- as.Date("2026-06-19")
chch_file$difference_days <- as.numeric(publish_date - chch_file$last_review) # create a new column "difference_days"
chch_file$difference_days

head(chch_file) # Check the new column difference_days exists
```
```{r review_days}
hist(chch_file$difference_days,
     main = "Distribution of Days since Last Review",
     xlab = "Days since last Review",
     ylab = "Listing counts",
     col = "blue")
```

Task - Find out which properties in New Zealand get the highest numbers of reviews (filter top 10%). How many of them are in Christchurch?